---
title: "NAA10 het liver"
author: "Andrew Garcia"
date: "2023-01-28"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# liver data
## stats and summary stats
```{r}
liver <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTLmgeuvv-9LZ7wmye300YHyft6hknpdF9volz23xEZENyhRra1DKxfLPytpTQnRd7s7LN2CgI5ZGsu/pub?gid=0&single=true&output=csv", header = T)

head(liver)
```
```{r}
liver$ID <-as.factor(liver$ID)
liver$Name <- as.factor(liver$Name)
liver$Genotype <- as.factor(liver$Genotype)
liver$Replicate <- as.factor (liver$Replicate)

head(liver)
```


```{r}
install.packages("Rmisc",repos = "http://cran.us.r-project.org")
```

```{r}
##Summary statistics for Normalized NAA10 signal, grouped by genotype and technical replicate
##Replicate grouping pertains to replicate blot grouping

library(Rmisc)
liver_sum <- summarySE(liver, measurevar="Normalized.Signal", groupvars=c("Genotype","Replicate"))

liver_sum
```

```{r}
##Summary statistics for Normalized NAA10 signal, grouped by genotype and technical replicate
##Replicate grouping pertains to replicate blot grouping

library(Rmisc)
liver_sum1 <- summarySE(liver, measurevar="Normalized.Signal", groupvars=c("Genotype"))

liver_sum1
```


```{r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
```



```{r}
#group means plus standard deviations come from summary stat dataframe
#essentially looking to plot this onto the individual data plot

library(ggplot2)
ggplot(liver_sum1, aes(x = Genotype, y = Normalized.Signal))+
  geom_point()+
  geom_errorbar(aes(ymin = Normalized.Signal - sd,
                ymax = Normalized.Signal + sd))
```

## plot with points and group means 
```{r}
#ggplot(heart) plots Normalized.Signal as a function of Genotype
#geom_point plots individual observations on this plot
#individual data points may need to be jittered
#geom_errorbar points to my group means df heart_sum1 and the bounds of the mean+/- standard error bar
#mean is plotted via stat_summary

ggplot(liver,aes(x = Genotype,
                 y = Normalized.Signal))+
  geom_point(aes(color = ID,
                 shape = Replicate),
             alpha = 1)+
  stat_summary(fun= mean, geom="point",
               shape=3, size=5, color="black", fill="black")+
  geom_errorbar(data = liver_sum1,
                aes(ymin = Normalized.Signal-sd,
                    ymax = Normalized.Signal+sd,
                    width = 0.25))+
  xlab("Naa10 genotype")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
  ggtitle("")
```


```{r}
ggplot(liver,aes(x = Genotype,
                 y = Normalized.Signal))+
  stat_summary(fun= mean, geom="point",
               shape=3, size=5, color="red", fill="black")+
  geom_errorbar(data = liver_sum1,
                aes(ymin = Normalized.Signal-sd,
                    ymax = Normalized.Signal+sd,
                    width = 0.25))+
    geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
  geom_point(aes(color = Name,
                 shape = Replicate),
             alpha = 1,
                 position = position_jitterdodge())+
  xlab("Naa10 genotype")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
  ggtitle("")
```



```{r}
ggplot(liver,aes(x = Genotype,
                 y = Normalized.Signal))+
  stat_summary(fun= mean, geom="point",
               shape=3, size=5, color="red", fill="black")+
  geom_errorbar(data = liver_sum1,
                aes(ymin = Normalized.Signal-sd,
                    ymax = Normalized.Signal+sd,
                    width = 0.25))+
  geom_point(aes(color = ID,
                 shape = Replicate),
             alpha = 1,
                 position = position_jitterdodge())+
  xlab("Naa10 genotype")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
  ggtitle("")
```


## jittered plot


```{r}
#same plot as above, but with jittered points
library(ggplot2)
liver.fig <- ggplot(liver,aes(x = Genotype,
                 y = Normalized.Signal))+
  geom_point(aes(color = ID,
                 shape = Replicate),
             alpha = 1,
             size = 2,
                 position = position_jitterdodge())+
  geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
  stat_summary(fun= mean, geom="point",
               shape=3, size=5,stroke = 1,color="red", fill="black")+
  geom_errorbar(data = liver_sum1,
                aes(ymin = Normalized.Signal-sd,
                    ymax = Normalized.Signal+sd,
                    width = 0.25))+
  xlab("Naa10 genotype")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
  theme_bw(base_size = 12, base_family = "Helvetica")

liver.fig
```



# hypothesis testing
## parametric
### two-sample t-test
null hypothesis: there is no mean difference in NAA10 signal between Naa10 female
genotypes

Alternative: there is a difference in NAA10 signal between Naa10 female genotypes
```{r}
##Unpaired t-test of NAA10 signal grouped by Genotype
t.test(liver$Normalized.Signal~liver$Genotype,
       paired = F,
       var.equal = F)
```

I fail to reject the null hypothesis

### differences between replicates
Null hypothesis: there is no difference in NAA10 signal of individual hearts on replicate blots
Alternative hypothesis: there is a difference in NAA10 signal of individual hearts on replicate blots


```{r}
install.packages("car",repos = "http://cran.us.r-project.org")
```

```{r}
library(car)
liver.aov <- lm(Normalized.Signal~Replicate, liver)
summary(liver.aov)
Anova(liver.aov,type = "III")

```

# #non-parametric testing
### NAA10 signal and Genotype
I have a big outlier in each genotype, so I will perform
non-parametric hypothesis tests.

```{r}
#Wilcox test instead of Welch's test
wilcox.test(Normalized.Signal~Genotype,liver)
```

p = 0.06253; I fail to reject the null hypothesis.


### Does replicate blot drive differences in NAA10 signal
```{r}
kruskal.test(Normalized.Signal~Replicate,liver)
```


### does median NAA10 signal differ?

```{r}
install.packages("coin",repos = "http://cran.us.r-project.org")
```

```{r}
library(coin)
mood_test(Normalized.Signal~Genotype,liver)
```


# earlier dataset with C57 females as inbred strain controls

```{r}
liver_c57 <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSOj04z27d0l76VXIjCAsvYw_0hn0_JUnmxJmXN6R0kuVZA3RytBDzfvBUv5WcNUg/pub?gid=788568614&single=true&output=csv",header = T)

head(liver_c57)
```

```{r}
liver_c57$Name <- as.factor(liver_c57$Name)
liver_c57$Genotype <- as.factor(liver_c57$Genotype)
liver_c57$Replicate <- as.factor (liver_c57$Replicate)

head(liver_c57)
```
```{r}
library(Rmisc)
liver_c57_sum <- summarySE(liver_c57, measurevar="Normalized.Signal", groupvars=c("Genotype"))

liver_c57_sum
```

```{r}
library(ggplot2)
```


```{r}
c57_fig <- ggplot(liver_c57,aes(x = Genotype,
                 y = Normalized.Signal))+
  geom_point(aes(shape = Replicate),
             alpha = 1.0,
             position = position_jitterdodge())+
  geom_boxplot(outlier.shape = NA,
               alpha = 0.1)+
  stat_summary(fun = mean, color = "black",
               position = position_dodge(0.75),width = 0.2,
             geom = "crossbar",
             show.legend = FALSE)+
  geom_errorbar(data = liver_c57_sum,
                aes(ymin = Normalized.Signal-sd,
                    ymax = Normalized.Signal+sd,
                    width = 0.2))+
  xlab("Naa10 genotype")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
  ggtitle("")

c57_fig
```


```{r}
##Plotted by biological replicate

ggplot(liver_c57, aes(x = Name,
                  y = Normalized.Signal,
                  shape = Replicate,
                  color = Genotype))+
         geom_point()+
  xlab("sample ID")+
  ylab("NAA10 signal normalized to
       REVERT total protein stain")+
   ggtitle("NAA10 levels in Naa10
          heterozygous mutant liver")

##verify genotypes of 4942 and 4924 (may be flipped in source data)
```

```{r}
t.test(liver_c57$Normalized.Signal~liver_c57$Genotype,
       paired = F,
       var.equal = F)
```
2-sample t-test indicates there is no significant difference in NAA10 signal
between genotypes; however, the means are non-equal.


```{r}
t.test(liver_c57$Normalized.Signal~liver_c57$Replicate,
       paired = T,
       var.equal = F)
```

Paired t-test indicates there is a significant difference between biological
replicates on replicate blots. First I will analyze based on Replicate.
Paired t-test already indicates there are differences between replicates so I
don't necessarily need to do post-hoc testing. However, I will use a
Tukey-Kramer
test.

If I include more variables in my model (Replicate, Genotype) that drive changes
in Normalized.Signal, I can use post-hoc testing to specify which
biological replicates differ.

For later reference: just re-run the samples again (ideally from new aliquots).

However, the samples for this dataset have been used up.

```{r}
liver_c57_aov <- aov(Normalized.Signal~Replicate,liver_c57)

summary(liver_c57_aov)
TukeyHSD(liver_c57_aov, conf.level = 0.95)
```
